\Large Clustering Trips Identified from Automatic Number \\ Plate Recognition Camera Scans

### 2018-02-22 classification.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays. We thus used all number-plate data collected between the 6th and 24th of February and exclude data from the 2 weekends in-between. Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$. Table~\ref{t:trips-tau} displays the number of trips, average trip degree and the proportion of trips of degree one, for varying values of $\Tau$. The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting. On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%. Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip. Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds. These were therefore ignored.

\begin{table}[t]
\centering
\tabcolsep=0.25cm
\begin{tabular}{c c c c}
  \hline
Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
  \hline
5 min & 13,603,759 & 1.46 & 0.70 \\
7.5 min & 12,394,709 & 1.60 & 0.64 \\
10 min & 11,690,791 & 1.69 & 0.61 \\
15 min & 10,823,333 & 1.83 & 0.57 \\
20 min & 10,305,860 & 1.92 & 0.54 \\
30 min &  9,653,499 & 2.04 & 0.52 \\
   \hline
\end{tabular}
\caption{Overview of trip data for varying values of $\Tau$.}
\label{t:trips-tau}
\vspace{-0.5cm}
\end{table}

@@

Table~\ref{t:features} depicts a sample of the resulting features vector. A total of 1034107 distinct vehicles were detected. However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated. We therefore, chose to remove three of the features represented in~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}. Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below zero) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642006 unique vehicles.

Clustering of vehicles was performed using the Hartigan and Wong \emph{k-means} algorithm, for each value of $\Tau$. The number of clusters \emph{k} was varied between 2 and 8 and executed with a maximum of 200 iterations and 100 different starting states of the algorithm. The Calinski-Harabasz criterion is used to pick the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}. The results of vehicle clustering are presented and discussed in the section below.

>>>>

Due to distinctly different traffic behaviour during weekends, we chose to consider only trips occurring during weekdays. Therefore, we used all number plate data collected between the 6th and 24th of February and excluded data from the 2 weekends in-between. Furthermore, as mentioned in section~\ref{s.trips}, we used fixed empirical values for $\Tau$. Table~\ref{t:trips-tau} displays the number of trips, average trip degree and the proportion of trips of degree one, for varying values of $\Tau$. The effect of incrementing $\Tau$ on resulting trips is clear: increased trip degrees and decreased trips containing just a single sighting. On the other hand, we did not set a value for $\tau$, but instead handled implausible trips by filtering all sightings with confidence below 85\%. Duplicates were filtered after identifying consecutive sightings by the same camera occurring within the same trip. Clock synchronisation errors were provided in milliseconds with none exceeding 5 seconds. These were hence ignored.

@@

Table~\ref{t:features} depicts a sample of the resulting features vector. A total of 1,034,107 distinct vehicles were detected. However, because there is a high percentage of trips containing a single sighting, some of these features were highly correlated. We therefore, chose to remove three of the features represented in table~\ref{t:features}: \emph{Average Sightings}, \emph{Average Distinct Routes} and \emph{Average Hour Difference}, to avoid the obfuscation of the natural clustering ~\cite{Kmeans}. Furthermore we considered that a trip of degree one has no destination (which explains values of average distinct destination below one) and we filtered all instances of vehicles where the total number of trips is lower than 3, resulting in 642,006 unique vehicles.

Clustering of vehicles was performed using the Hartigan and Wong \emph{k-means} algorithm, for each value of $\Tau$. The number of clusters \emph{k} was varied between 2 and 8 and executed with a maximum of 200 iterations and 100 different starting states of the algorithm. The Calinski-Harabasz criterion is used to determine the best value of $k$, the one that minimises the within-cluster and between-cluster errors and provides the more natural clusters ~\cite{Kmeans}. The results of trip clustering are presented and discussed in the next section.

### 2018-02-22 discussion.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

Most urban cities in the world possess a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections. We also identify two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips. Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network. Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

>>>>

Most urban cities in the world employ a network of ANPR cameras that are used for law enforcement as well as traffic monitoring and control. Number plate data collected in the Tyne and Wear area is stored and leveraged by the UTMC for computing average journey times across a selection of sensitive roads. However, number plate data could be used more extensively to identify and study individual and collective travel patterns. In this paper, we have presented a set of definitions and constraints that establish a conceptual foundation for identifying vehicle trips from number plate detections.

We have also identified two parameters, $\tau$ and $\Tau$ as critical in the discrimination of plausible and implausible trips. Hence, future work should first and foremost focus on developing formal methods to estimate these parameters from observed distributions of travel times and by applying knowledge about the structure of the road network. Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and be amplified in posterior analysis done using trip data. Moreover, methods for addressing issues concerning camera performance, namely wrong and duplicate scans, should be further developed and researched.

### 2018-02-22 trips.tex
Made some text changes and some re-placement of tables/figures: discussion/errors in plate scanning

The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes. However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed. Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted. Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times. The same rationale can be applied to $\tau$.

@@

\begin{figure*}[!ht]%
  \centering
  \begin{subfigure}[c]{.5\textwidth}
    \small
    \tabcolsep=0.09cm
    \begin{tabular}{c c c c c c c}
      \hline
      Vehicle & Camera & Timestamp & Trip & Sighting & \thead{Journey \\Time} & \thead{Trip \\Id}\\
      \hline
      2362920 & 1014 & \makecell{2017-02-01 \\ 00:00:06} &   1 &   1 & NA & 21 \\
      2362920 & 1044 & \makecell{2017-02-01 \\ 00:01:28} &   1 &   2 & 82.38 & 21 \\
      2362920 &  35 & \makecell{2017-02-01 \\ 00:02:32} &   1 &   3 & 63.50 & 21\\
      2362920 &  32 & \makecell{2017-02-01 \\ 00:04:38} &   1 &   4 & 125.95 & 21\\
       \hline
    \end{tabular}
    \label{fig:trip-example-table}
  \end{subfigure}\hfill
  %\qquad
  \begin{subfigure}[c]{.48\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{trip-example.pdf}
    \label{fig:trip-example-map}
  \end{subfigure}\hfill
  \caption{Example of a trip of degree 4. On the left side the corresponding data table is shown. The $u_{th}$ trip of each vehicle is given by the variable \emph{Trip}, whereas the $i_{th}$ sighting is given by variable \emph{Sighting}. The variable \emph{JourneyTime} gives the travel time from the previous to current sighting. Lastly, the variable \emph{TripId} represents the unique sequence of cameras that describes the trip. This allows trips to be grouped and summarised not only in terms of their origins and destination, but also routes. On the right side, the same trip is plotted on a map. The circles represent camera locations whereas the lines represent the fastest driving routes between sightings rather than the true route taken by the vehicle. Even though no routing information is available for each consecutive pair of sightings, the observed journey times can be compared against the distribution of collective journey times to rank the set of most likely routes chosen (which can be determined for instance by Stochastic User Equilibrium~\cite{Castillo2008}).}%
  \label{fig:trip-example}%
\end{figure*}

On the other hand, a consistent set of models and methodologies for estimating $\Tau$ and $\tau$, under a variety of circumstances and road anomalies, remains an open research problem. Thus, for the purposes of this work, we consider these parameters to be fixed. Nevertheless, the importance of estimating these parameters should not be understated. Any errors in trip identification that occur due to poor estimation and filtering methods will propagate and will be amplified in posterior analysis done using trip data. To briefly exemplify this, figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values ($\tau$ is ignored).

\subsubsection*{Duplicate scannings}

@@

 Although the estimation of $\gamma$ carries similar considerations and consequences as those of estimating $\Tau$ and $\tau$, most duplicates can be identified in consecutive sightings of the same camera within the same trip. Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.

\subsubsection*{Errors in plate scanning}

Number plate recognition cameras have accuracy rates of 99.9\% or higher. If we consider that on average 1 to 10 out of 10000 scans are misclassified number-plates then approximately 200 to 2000 scans everyday are incorrect. We categorise the errors made by ANPR cameras into two types:
\begin{enumerate*}[label=(\roman*)]
  \item a passing vehicle is not detected by the camera;
  \item a different vehicle is detected instead.
\end{enumerate*}
In the first case, there is no observed data. The corresponding trip sequence for the affected vehicle will be missing a sighting but we will have no indication of this. It should be noted that this could also happen in the case where a vehicle does not take the `normal' route and pass this intermediary camera. For the second case, there is a recorded sighting, but this will be assigned to the incorrect vehicle. Even though, the camera provides a value of \emph{confidence} which helps in diagnosing incorrect scans, errors still occur for high values of confidence. Even so, in this work we do not present a complex solution for this issue.

>>>>

The simplest approach to choosing the value of $\Tau$ is to pick a fixed empirical value, such as 5 or 10 minutes. However, if the distance between two cameras is greater than another pair of cameras, then it makes sense that $\Tau$ is relaxed. Similarly, if there is an anomaly in the road network, such as a traffic jam, and the routes connecting the two cameras are affected, then the value of $\Tau$ should also be adapted. Hence, $\Tau$ should be a function of the distance between the two cameras (or, more accurately, of the top n-routes between these) and the distribution of observed journey times. The same rationale can be applied to $\tau$. However, due to time constraints, in this work we simply used several fixed values of $\Tau$ and set $\tau$ to zero. Figure~\ref{fig:length-dist} shows how the number of trips per degree of trip varies by fixing $\Tau$ at different empirical values.

@@

\begin{figure*}[!ht]%
  \centering
  \begin{subfigure}[c]{.5\textwidth}
    \small
    \tabcolsep=0.09cm
    \begin{tabular}{c c c c c c c}
      \hline
      Vehicle & Camera & Timestamp & Trip & Sighting & \thead{Journey \\Time} & \thead{Trip \\Id}\\
      \hline
      2362920 & 1014 & \makecell{2017-02-01 \\ 00:00:06} &   1 &   1 & NA & 21 \\
      2362920 & 1044 & \makecell{2017-02-01 \\ 00:01:28} &   1 &   2 & 82.38 & 21 \\
      2362920 &  35 & \makecell{2017-02-01 \\ 00:02:32} &   1 &   3 & 63.50 & 21\\
      2362920 &  32 & \makecell{2017-02-01 \\ 00:04:38} &   1 &   4 & 125.95 & 21\\
       \hline
    \end{tabular}
    \label{fig:trip-example-table}
  \end{subfigure}\hfill
  %\qquad
  \begin{subfigure}[c]{.48\textwidth}
    \centering
    \includegraphics[width=1\linewidth]{trip-example.pdf}
    \label{fig:trip-example-map}
  \end{subfigure}\hfill
  \caption{Example of a trip of degree 4. On the left side the corresponding data table is shown. The $u_{th}$ trip of each vehicle is given by the variable \emph{Trip}, whereas the $i_{th}$ sighting is given by variable \emph{Sighting}. The variable \emph{JourneyTime} gives the travel time from the previous to current sighting. Lastly, the variable \emph{TripId} represents the unique sequence of cameras that describes the trip. This allows trips to be grouped and summarised not only in terms of their origins and destination, but also routes. On the right side, the same trip is plotted on a map. The circles represent camera locations whereas the lines represent the fastest driving routes between sightings rather than the true route taken by the vehicle. Even though no routing information is available for each consecutive pair of sightings, the observed journey times can be compared against the distribution of collective journey times to rank the set of most likely routes chosen (which can be determined for instance by Stochastic User Equilibrium~\cite{Castillo2008}).}%
  \label{fig:trip-example}%
  \vspace{-0.40cm}
\end{figure*}
\vspace{-0.25cm}

\subsection{Duplicate scannings}

@@

 Although the estimation of $\gamma$ carries similar considerations and consequences as those of estimating $\Tau$ and $\tau$, most duplicates can be identified in consecutive sightings of the same camera within the same trip. Even though a poor estimation of $\gamma$ also has an impact on error propagation, this decreases substantially after filtering duplicates according to the heuristic above, due to the low occurrence of cycles in trips.
\vspace{-0.25cm}

\subsection{Errors in plate scanning}

ANPR cameras have average accuracy rates of 99.9\% and higher. If we consider that on average 1 to 10 out of 10000 number plate scans are misclassified number-plates then approximately between 200 to 2000 scans everyday are incorrect. These errors propagate and lead to innacuracies in the identification of trips. In fact, misclassifications affect the trip sequences of two vehicles: the true passing vehicle and the vehicle erroneously detected instead. The true vehicle will be missing a sighting in the correspoding trip sequence vector whereas the other vehicle's trip sequence will contain an extra invalid sighting. Moreover, ANPR cameras may exceptionally fail to detect passing vehicles. Even though only the passing vehicle is affected in this case, it's trip sequence vector is nonetheless affected. Therefore, due to their impact in trip identification, it is important to detect and address missing and misclassified scans.

We filer invalid sightings by removing all sightings containing a \emph{confidence} value below $85\%$. However, we do not address missing scans in this work.

\begin{table}[!ht]
\centering
\tabcolsep=0.25cm
\begin{tabular}{c c c c}
  \hline
Tau & Trips & \thead{Average\\Degree} & \thead{Proportion of Trips \\ Degree 1} \\
  \hline
5 min & 13,603,759 & 1.46 & 0.70 \\
7.5 min & 12,394,709 & 1.60 & 0.64 \\
10 min & 11,690,791 & 1.69 & 0.61 \\
15 min & 10,823,333 & 1.83 & 0.57 \\
20 min & 10,305,860 & 1.92 & 0.54 \\
30 min &  9,653,499 & 2.04 & 0.52 \\
   \hline
\end{tabular}
\caption{Overview of trip data for varying values of $\Tau$.}
\label{t:trips-tau}
\vspace{-0.5cm}
\end{table}
